---
title: "Classification Flexible Discriminant Analysis. Iris Data"
author: "Jane Kathambi"
date: "8 June 2018"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---
# Introduction
A classification problem involves predicting a non-numerical value—that is, a categorical variable, also known as discrete variable.

Most of the classification algorithms computes the probability of belonging to a given class. Observations are then assigned to the class that have the highest probability score.

Generally, you need to decide a probability cutoff above which you consider the an observation as belonging to a given class.

## Recap
** separable data:** 
* LDA: Small training set. Also Linear kernel svms.
* QDA: Large training set. Also Linear kernel svms.
* RDA: Large training set and too many features. Also Linear kernel svms.

**inseparable data:**
* FDA: Multivariate inseparable data sets. Also radial kernel svms.
* MDA: Classes have sub-classes which are distributed. Also radial kernel svms.

#The Iris data set
The iris data set will be used for multiclass classification tasks. It contains the length and width of sepals and petals for three iris species. We want to predict the species based on the sepal and petal parameters.

This data has three classes of which two are inseparable classes as we will see shortly. So FDA will model this data well. Radial kernel svm and MDA will model this data well too.

# Classification:  FDA (Flexible discriminant analysis)
FDA is a flexible extension of LDA that uses non-linear combinations of predictors such as splines. FDA is useful to model multivariate non-normality or non-linear relationships among variables within each group, allowing for a more accurate classification.

# Load the required libraries.
* tidyverse. for easy data manipulation and visualization
* caret. for easy machine learning workflow. createDataPartition for partitioning the data into test set and train set.
* fda library: has the fda function.

```{r}
library(tidyverse)
library(caret)
library(fda)
```

# Exploring the Iris Data

```{r}
# Load the data
data("iris")

# check for Nas
anyNA(iris)

# Inspect the data
sample_n(iris, 3)

# dim
dim(iris)

#internl structure
glimpse(iris)

# Covariance scatterplot Matrix. View how each variable varies with the rest as well as how the classes are distributed
pairs(iris, col=iris$Species)

```

From the covariance matrix above it is evident that there are three classes two of which are inseparable but one is separable from the other two classes. So FDA will model this data well. Radial kernel svm and MDA will model this data well too.

# Normalize the data. Categorical variables are automatically ignored.
```{r}
# Estimate preprocessing parameters
preproc.param <- iris %>% 
  preProcess(method = c("center", "scale"))

# Transform the data using the estimated parameters
standardized_iris <- preproc.param %>% predict(iris)

```

# Split the data into training and test set
```{r}
set.seed(123)
training.samples <- standardized_iris$Species %>% 
  createDataPartition(p = 0.8, list = FALSE)

train.data  <- standardized_iris[training.samples, ]
test.data <- standardized_iris[-training.samples, ]
```

# Fit the FDA model
```{r}
# Fit the FDA model
fda_model <- fda(Species~., data = train.data)
fda_model
```

# Plot the model on classification of train.data (predictions on train data)
```{r}
# plot the model
plot(fda_model, train.data)
```

# Plot the model on classification of test.data (predictions on new data)
```{r}
# plot the model
plot(fda_model, test.data)
```

# FDA Model accuracy
For fda model the output of predictions is predicted classes. So we access model accuracy as follows:
mean(preds == test.data$Species)
```{r}
# Make predictions
preds <- fda_model %>% predict(test.data)
# Model accuracy
mean(preds == test.data$Species)


```

Our model has an accuracy of 1. This implies that the model has done the best that there is to do. 100% correct classifications. 

The best models for this data are FDA and MDA.

# Model Accuracies Ranking
The following is a ranking of how the models perfomed on classifying this iris data starting from the best model. 
1. Flexible Discriminant Analysis: 1 accuracy
1. Mixture Discriminant Analysis: 1 accuracy
2. Radial Kernel SVM classifier:  0.967 accuracy

