---
title: "UI Inputs, checkbox input"
author: "Jane Kathambi"
date: "2 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
runtime: shiny
---

# Introduction
We already have a predefined sidebarLayout. This layout has two drop down menus defined for the x and y input and one output which is a scatter plot.
We also have a predefined server functions which plots out/renders the scatterplot based on the user input.

We are going to add a checkbox input that specifies whether the plotted data should be shown in a data table. We will add the following:
1. UI: Add an input widget that the user can interact with to check/uncheck the box
2. UI: Add an output defining where the data table should appear.
3. Server: Add a reactive expression that creates the data table if the checkbox is checked


# Load the required libraries
We load DT library in addition to shiny, ggplot2, dplyr libraries
```{r}
library(shiny)
library(dplyr)
library(ggplot2)
library(DT)
```

#load the movies data set

We load the data before the UI and Server defenitions so that it can be used in both

```{r}
movies<- get(load("data/movies.RData") )
head(movies)
```

# Initializing the shiny app

## UI definition
We will:
1. UI: Add an input widget that the user can interact with to check/uncheck the box
2. UI: Add an output defining where the data table should appear.

```{r}
#User Interface. Controls the layout and appearance of App
#Define UI for application that plots features of movies 
ui<-fluidPage(
  

  # Sidebar layout with a input and output definitions 
  sidebarLayout(
    
   # Inputs
    sidebarPanel( 
      
            # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("IMDB rating" = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score" = "critics_score", 
                              "Audience score" = "audience_score", 
                              "Runtime" = "runtime"), 
                  selected = "audience_score"),
      
      
      # Select variable for x-axis
     selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("IMDB rating" = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score" = "critics_score", 
                              "Audience score" = "audience_score", 
                              "Runtime" = "runtime"), 
                  selected = "critics_score"),
      
      # Add a checkbox widget
     checkboxInput(inputId = "show_data", 
                  label = "Show Data Table:",
                  value=TRUE #TRUE means that the box is initially checked
                  )
    
    ),# end of sidebar panel
    
   # Outputs
    mainPanel(
      
      # show scatterplot
      plotOutput(outputId = "scatterplot"),
      
      #show data table
      DT::dataTableOutput(outputId = "moviestable")
      
    )# end of mainPanel
   
  )# end of sidebarLayout
  
)# end of fluidPage
```


## Server definition
We will:
3. Server: Add a reactive expression that creates the data table if the checkbox is checked
```{r}
# Server Fnction. Contains instructions needed to build app
# Define server function required to create the scatterplot
server <- function(input, output) {

  # Create scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y)) +
      geom_point(alpha=0.4)
  })
  
  # Print data table if checked
  output$moviestable<-DT::renderDataTable(
    {
      if(input$show_data)
      {
        DT::datatable(
          data = movies%>%select(1:7), #select the first up to the seventh column
          options = list(pageLength=50),# number of rows per window
          rownames = FALSE #do not show row names i.e numbering
          
        )
      }
    }
  )
}

```

## Creating the Shiny app

```{r}
# Creates the shiny app object
shinyApp(ui=ui, server=server)
```

