---
title: "UI Inputs, numeric input"
author: "Jane Kathambi"
date: "2 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
runtime: shiny
---

# Introduction
We are going to show a sample of the movies data as a table. The sample size is determined by the users input. We will use the renderDataTable function from the DT package to render the table.

We are going to add a numericInput that specifies the sample size. We will add the following:
1. UI: Add an numericInput widget that the user can interact with to select the sample size.
2. UI: Add an output defining where the data table should appear.
3. Server: Add a reactive expression that creates a sample of the data table based on the user input of the sample size.


# Load the required libraries
We load DT library in addition to shiny, ggplot2, dplyr libraries
```{r}
library(shiny)
library(dplyr)
library(ggplot2)
library(DT)
```

#load the movies data set

We load the data before the UI and Server defenitions so that it can be used in both

```{r}
movies<- get(load("data/movies.RData") )
head(movies)

```

# Calculate total number of movies

We calculate n_total (total number of movies in the dataset) as nrow(movies) before the UI definition. We willuse this total when creating the numeric input widget as well as to render the table.

```{r}
#Calculate n_total (total number of movies in the dataset)
n_total <- nrow(movies)
```

# Initializing the shiny app

## UI definition
We will:
1. UI: Add an numericInput widget that the user can interact with to select the sample size.
2. UI: Add an output defining where the data table should appear.

```{r}
#User Interface. Controls the layout and appearance of App
#Define UI for application that plots features of movies 
ui<-fluidPage(
  

  # Sidebar layout with a input and output definitions 
  sidebarLayout(
    
   # Inputs
    sidebarPanel( 
      
      # Text instructions
      HTML(paste("Enter a value between 1 and", n_total)),
      
      # Numeric input for sample size
      numericInput(inputId = "n",
                   label = "Sample size:",
                   min=1,
                   max=n_total,
                   value = 30,
                   step = 1)
    
    ),# end of sidebar panel
    
    # Output: Show data table
    mainPanel(
      DT::dataTableOutput(outputId = "moviestable")
    )# end of mainPanel
   
  )# end of sidebarLayout
  
)# end of fluidPage
```


## Server definition
We will:
3. Server: Add a reactive expression that creates a sample of the data table based on the user input of the sample size.

Run the code and then in the Shiny app (not in the code), delete the numeric value (30) in the sample size box. You will encounter the error: 
* Error: size is not a numeric or integer vector.

In order to avoid such errors, which users of your app could very easily encounter, we need to hold back the output from being calculated if the input is missing. The req function is the simplest and best way to do this, it ensures that values are available ("truthy") before proceeding with a calculation or action. If any of the given values is not truthy, the operation is stopped by raising a "silent" exception (neither logged by Shiny, nor displayed in the Shiny app's UI).

In the server: Inside the renderDataTable function, add req(input$n) before movies_sample is calculated.

```{r}
# Define server function required to create the scatterplot
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
    req(input$n)
    movies_sample <- movies %>% #we want to only show a sample based on the n input.
      sample_n(input$n) %>%# apply sample_n function to n
      select(title:studio)# select title column up tp studio column
    
    #Then render the movies_sample using DT::datatable() function
    DT::datatable(data = movies_sample, 
                  options = list(pageLength = 10), #10 rows per window
                  rownames = FALSE)# dont show row counts/names.
  })
  
}

```

## Creating the Shiny app

```{r}
# Creates the shiny app object
shinyApp(ui=ui, server=server)
```

