---
title: "UI Inputs, Selectize for multiple selections"
author: "Jane Kathambi"
date: "2 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
runtime: shiny
---

# Select to selectize

How can we select multiple studios to display all at once? We can use:
* selectize. Selectize suggest names of studios as you type them. 
    + selectizeInput(inputId, ..., options = NULL, width = NULL)
* selectInput and set multiple to true. Selectize in selectInput is set to true by default.
    + selectInput(inputId, label, choices, selected = NULL, multiple = FALSE,
  selectize = TRUE, width = NULL, size = NULL)


We will use selectInput and set multiple to true. Selectize in selectInput is set to true by default.

There are 211 unique studios represented in this dataset, we need a better way to select than to scroll through such a long list, and we address that with the selectize option, which will suggest names of studios as you type them.

# Load the required libraries
We load DT library in addition to shiny, ggplot2, dplyr libraries
```{r}
library(shiny)
library(dplyr)
library(ggplot2)
library(DT)
```

#load the movies data set

We load the data before the UI and Server defenitions so that it can be used in both

```{r}
movies<- get(load("data/movies.RData") )
head(movies)

```

# Obtain a list of all the unique studios

Let us obtain a sorted list of all the unique studios.

```{r}
#a sorted list of all the unique studios
all_studios <- sort(unique(movies$studio))

```

# Initializing the shiny app

## UI definition

```{r}
#User Interface. Controls the layout and appearance of App
#Define UI for application that plots features of movies 
ui<-fluidPage(
  

  # Sidebar layout with a input and output definitions 
  sidebarLayout(
    
   # Inputs
    sidebarPanel( 
      
      #select multiple studios
      selectInput(inputId = "studio",
                  label = "Select studio:",
                  choices = all_studios,
                  selected = "20th Century Fox",
                   multiple = TRUE)
    
    ),# end of sidebar panel
    
    # Output: Show data table
    mainPanel(
      DT::dataTableOutput(outputId = "moviestable")
    )# end of mainPanel
   
  )# end of sidebarLayout
  
)# end of fluidPage
```

## Server definition

1. Add req(input$n). It prevents an error from being displayed when no studio input is provided.

2. Let the filter() to use the logical operator %in% (instead of ==) so that it works when multiple studios are selected. i.e filter(studio %in% input$studio) %>%

```{r}
# Server
server <- function(input, output) {
  
  # Create data table
  output$moviestable <- DT::renderDataTable({
  req(input$studio)
    movies_from_selected_studios <- movies %>%
      filter(studio %in% input$studio) %>%
      select(title:studio)
    DT::datatable(data = movies_from_selected_studios, 
                  options = list(pageLength = 10), 
                  rownames = FALSE)
  })
  
}


```

## Creating the Shiny app

```{r}
# Creates the shiny app object
shinyApp(ui=ui, server=server)
```

