---
title: "UI Inputs, Date Range"
author: "Jane Kathambi"
date: "2 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
runtime: shiny
---

# Date Range
The date range function:

dateRangeInput(inputId, label, start = NULL, end = NULL, min = NULL,
  max = NULL, format = "yyyy-mm-dd", startview = "year", weekstart = 0,
  language = "en", separator = " to ", width = NULL, autoclose = TRUE)

# Load the required libraries
We load DT library in addition to shiny, ggplot2, dplyr libraries
```{r}
library(shiny)
library(dplyr)
library(ggplot2)
```

#load the movies data set

We load the data before the UI and Server defenitions so that it can be used in both

```{r}
movies<- get(load("data/movies.RData") )
head(movies)

```

# Obtain the min and max dates from the data

These will be used in the dateRangeInput widget.
```{r}
min_date <- min(movies$thtr_rel_date)
max_date <- max(movies$thtr_rel_date)
```

# Initializing the shiny app

## UI definition

```{r}
# UI
ui <- fluidPage(
    sidebarLayout(
    
    # Input(s)
    sidebarPanel(
      
      # Explanatory text
      HTML(paste0("Movies released between the following dates will be plotted. 
                  Pick dates between ", min_date, " and ", max_date, ".")),
      
      # Break for visual separation
      br(), br(),
      
      # Date input
      dateRangeInput(inputId = "date", # date is a vector of two dates i.e start date and end date.  
                label = "Select dates:",
                start= "2013-01-01",
                end= "2014-01-01",
                startview="year",
                min = min_date, max = max_date)
    ),
    
    # Output(s)
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)
```

## Server definition

```{r}
# Server
server <- function(input, output) {
  
  # Create the plot
  output$scatterplot <- renderPlot({
    req(input$date)
    movies_selected_date <- movies %>%
      mutate(thtr_rel_date = as.Date(thtr_rel_date)) %>% # convert thtr_rel_date to Date format
      filter(thtr_rel_date >= input$date[1] & thtr_rel_date <= input$date[2])
    ggplot(data = movies_selected_date, aes(x = critics_score, y = audience_score, color = mpaa_rating)) +
      geom_point()
  })
  
}


```

## Creating the Shiny app

```{r}
# Creates the shiny app object
shinyApp(ui=ui, server=server)
```

