---
title: "Introduction to Building Web Apps with Shiny"
author: "Jane Kathambi"
date: "2 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
runtime: shiny
---

# Introduction

# Load the required libraries

```{r}
library(shiny)
library(dplyr)
library(ggplot2)
library(rsconnect)
```

#load the movies data set

We load the data before the UI and Server defenitions so that it can be used in both

```{r}
movies<- get(load("data/movies.RData") )
head(movies)
```

# Initializing the shiny app

## UI definition
With UI definition we do the following:

* Inputs are defined and laid out. They are defined in the **side bar panel** of the sidebar Layout i.e side bar panel contains input controls.

    + Our **side bar panel** contains three dropdown menus (x,y,color) created with the selectInput widget function.  The selectInput widget is used to allow the user to select which variables should be plotted on the x and y axes of the scatterplot. The selectInput function has four arguments:
    
        - An **inputId** is the input value that the app will internally use to access the value selected by the user when building a scatterplot.
        - A **label** that is displayed in the app.
        - A list of **choices** to pick from. Note that choices takes a named vector, and the name rather than the value (which must match variable names in the data frame) is displayed to the user.
        - A **selected** choice for when the app first launches.
        * Outputs are laid out. They are defined in the **main Panel** of the side bar Layout, i.e main panel contains the output.

    + The **main panel** contains the output elements that get created in the server function. In this case, the main panel contains only one component which is a plot output.

```{r}
#User Interface. Controls the layout and appearance of App
#Define UI for application that plots features of movies 
ui<-fluidPage(
  

  # Sidebar layout with a input and output definitions 
  sidebarLayout(
    
   # Inputs
    sidebarPanel( 
      
            # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("IMDB rating" = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score" = "critics_score", 
                              "Audience score" = "audience_score", 
                              "Runtime" = "runtime"), 
                  selected = "audience_score"),
      
      
      # Select variable for x-axis
     selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("IMDB rating" = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score" = "critics_score", 
                              "Audience score" = "audience_score", 
                              "Runtime" = "runtime"), 
                  selected = "critics_score"),
     
    # select variable for color
      selectInput(
      inputId="z",
      label="Color by:",
      choices=c("Title type"="title_type", "Genre"="genre", "MPAA rating"="mpaa_rating", "Critics rating"="critics_rating", "Audience rating"="audience_rating"),
      selected="mpaa_rating"
      )
    
    ),# end of sidebar panel
    
   # Outputs
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )# end of mainPanel
   
  )# end of sidebarLayout
  
)# end of fluidPage
```


## Server definition
Defines the relationship between inputs and outputs. 

The **server function** takes three arguments that is input, output and session. Session is used by the application to assess session specific information and functions. For now we will not deal with session.

The **renderPlot function** specifies how the **plotOutput** should be updated.

* Outputs are calculated.
* Any other calculations needed for outputs are perfomed.

There are three major rules of creating server functions:

* Save objects to display to output$xx in this case output$scatterplot
* Build objects to display with render*() in this case renderPlot({}). 
  Reactive outputs should be built with render*() functions. Reactive output meas that output changes with change in input.
* Use input values with input$xx in this case input$x, input$y and input$z

There are different render fuctions which include:

* renderTable() 
* DT::renderDataTable()
* renderImage() 
* renderText()
* renderPrint()
* renderUI()

```{r}
# Server Fnction. Contains instructions needed to build app
# Define server function required to create the scatterplot
server <- function(input, output) {

  # Create scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y, color = input$z)) +
      geom_point()
  })
}

```

## Creating the Shiny app

```{r}
# Creates the shiny app object
shinyApp(ui=ui, server=server)
```

