---
title: "Handling XML"
author: "Jane Kathambi"
date: "11 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Overview

Sometimes data is a TSV or nice plaintext output. Sometimes it's XML and/or JSON. 

This chapter walks you through what XML is, how to convert xml into R-like objects, and how to extract data from them. 

You'll practice by examining the revision history for a Wikipedia article retrieved from the Wikipedia API using httr and xml2.

An API returns various data formats but the two most common data formats are:
1. xml
2. json

Lets check out xml.

# xml


# Load the required libraries

```{r}
library(xml2)

```

# Examining XML documents
Just like JSON, you should first verify the response is indeed XML with http_type() and by examining the result of content(r, as = "text"). Then you can turn the response into an XML document object with read_xml().

One benefit of using the XML document object is the available functions that help you explore and manipulate the document. For example xml_structure() will print a representation of the XML document that emphasizes the hierarchical structure by displaying the elements without the data.

In this exercise you'll grab the same revision history you've been working with as XML, and take a look at it with xml_structure().

Let us begin by writing a function to get the revision history for the Wikipedia article for "Hadley Wickham".

```{r}
rev_history <- function(title, format = "json"){
  
  if (title != "Hadley Wickham") {
    stop('rev_history() only works for `title = "Hadley Wickham"`')
  }
  
  if (format == "json"){
    resp <- readRDS("had_rev_json.rds")
  } 
  
  else if (format == "xml"){
    resp <- readRDS("had_rev_xml.rds")
  } 
  
  else {
    stop('Invalid format supplied, try "json" or "xml"')
  }
  
  resp  
}

```


Let us use our function to obtain the revision history for "Hadley Wickham"

Get the XML version of the revision history for the Wikipedia article for "Hadley Wickham", by calling rev_history("Hadley Wickham", format = "xml"), store it in resp_xml.

```{r}
# Get XML revision history
resp_xml <- rev_history("Hadley Wickham", format = "xml")
```

Check the response type of resp_xml to confirm the API returned an XML object.
```{r}
# Check response is XML 
http_type(resp_xml)
```

You can't always trust a header, so check the content looks like XML by calling content() on resp_xml with as = "text", store in rev_text.

```{r}
# Examine returned text with content()
rev_text <- content(resp_xml, as="text")
rev_text
```

Turn rev_text into an XML object with read_xml() from the xml2 package, store as rev_xml.

```{r}
# Turn rev_text into an XML document
rev_xml <- read_xml(rev_text)
```

Call xml_structure() on rev_xml to see the structure of the returned XML. Can you see where the revisions are?

```{r}
# Examine the structure of rev_xml
xml_structure(rev_xml)
```

xml_structure() helps you understand the structure of your document, without overwhelming you with content.

# Extracting XML data
XPATHs are designed to specifying nodes in an XML document. Remember /node_name specifies nodes at the current level that have the tag node_name, where as //node_name specifies nodes at any level below the current level that have the tag node_name.

xml2 provides the function xml_find_all() to extract nodes that match a given XPATH. For example, xml_find_all(rev_xml, "/api") will find all the nodes at the top level of the rev_xml document that have the tag api. Try running that in the console. You'll get a nodeset of one node because there is only one node that satisfies that XPATH.

The object returned from xml_find_all() is a nodeset (think of it like a list of nodes). To actually get data out of the nodes in the nodeset, you'll have to explicitly ask for it with xml_text() (or xml_double() or xml_integer()).

Use what you know about the location of the revisions data in the returned XML document extract just the content of the revision.

Before doing anything first look at the structure of the xml documet using xml_structure(rev_xml)
```{r}
xml_structure(rev_xml)
```

Use xml_find_all() on rev_xml to find all the nodes that describe revisions by using the XPATH, "/api/query/pages/page/revisions/rev".

```{r}
# Find all nodes using XPATH "/api/query/pages/page/revisions/rev"
xml_find_all(rev_xml, "/api/query/pages/page/revisions/rev")
```


Use xml_find_all() on rev_xml to find all the nodes that are in a rev node anywhere in the document, store in rev_nodes.

```{r}
rev_nodes<-xml_find_all(rev_xml, "//rev")
```

Extract the contents from each node in rev_nodes, by passing rev_nodes to xml_text().

```{r}
# Use xml_text() to get text from rev_nodes
xml_text(rev_nodes)
```

# Extracting XML attributes
Not all the useful data will be in the content of a node, some might also be in the attributes of a node. 

To extract attributes from a nodeset, xml2 provides xml_attrs() and xml_attr().

xml_attrs() takes a nodeset and returns all of the attributes for every node in the nodeset. 

xml_attr() takes a nodeset and an additional argument attr to extract a single named argument from each node in the nodeset.

In this exercise you'll grab the user and anon attributes for each revision.

You'll see xml_find_first() in the sample code. It works just like xml_find_all() but it only extracts the first node it finds.

Extract rev_nodes and first_rev_node in the document.
```{r}
#Extract rev_nodes
rev_nodes=xml_find_all(rev_xml,"//rev")

first_rev_node=xml_find_first(rev_xml,"//rev")
```


Now it is time to explore the difference between xml_attrs() and xml_attr().

Use xml_attrs() on first_rev_node to see all the attributes of the first revision node.

```{r}
# Find all attributes with xml_attrs() for first_rev_node
xml_attrs(first_rev_node)
```

Use xml_attr() on first_rev_node along with an appropriate attr argument to extract the user attribute from the first revision node.

```{r}
# Find user attribute with xml_attr() for first_rev_node
xml_attr(first_rev_node, attr='user')
```

Now use xml_attr() again, but this time on rev_nodes to extract the user attribute from all revision nodes.

```{r}
# Find user attribute for all rev nodes
xml_attr(rev_nodes, attr='user')
```

Use xml_attr() on rev_nodes to extract the anon attribute from all revision nodes.
```{r}
# Find anon attribute for all rev nodes
xml_attr(rev_nodes, attr='anon')
```

If a node didn't have the anon attribute xml_attr() returned an NA





















