---
title: "Wrap up"
author: "Jane Kathambi"
date: "11 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Wrapup: returning nice API output
How might all this work together? A useful API function will retrieve results from an API and return them in a useful form. 

We have already seen how to write a function that retrieves data from an API that relied on content() to convert it to a useful form. To write a more robust API function you shouldn't rely on content() but instead parse the data yourself.

To finish up this chapter you'll do exactly that: write get_revision_history() which retrieves the XML data for the revision history of page on Wikipedia, parses it, and returns it in a nice data frame.

We will call get_revision() function inside of our new function to return the response that we will parse.

You can assume this function returns the raw response and follows the best practices you learnt in Chapter 2, like using a user agent, and checking the response status.


# Load the required libraries

```{r}
library(xml2)
```

# Create rev_history() function to obtain raw repsonse

Let us begin by writing a function to get the revision history for the Wikipedia article for "Hadley Wickham".

```{r}
rev_history <- function(title, format = "json"){
  
  if (title != "Hadley Wickham") {
    stop('rev_history() only works for `title = "Hadley Wickham"`')
  }
  
  if (format == "json"){
    resp <- readRDS("had_rev_json.rds")
  } 
  
  else if (format == "xml"){
    resp <- readRDS("had_rev_xml.rds")
  } 
  
  else {
    stop('Invalid format supplied, try "json" or "xml"')
  }
  
  resp  
}

```

# Create get_revision_history() function to parse response

We will call the get_revision function to obtain the response then parse it and finally, call get_revision_history() with article_title = "Hadley Wickham".

```{r}

get_revision_history <- function(article_title){
  # Get raw revision response
  rev_resp <- rev_history(article_title, format = "xml")
  
  # Turn the content() of rev_resp into XML
  rev_xml <- read_xml(content(rev_resp, "text"))
  
  # Find revision nodes
  rev_nodes <- xml_find_all(rev_xml, "//rev")

  # Parse out usernames
  user <- xml_attr(rev_nodes, attr='user')
  
  # Parse out timestamps
  timestamp <- readr::parse_datetime(xml_attr(rev_nodes, "timestamp"))
  
  # Parse out content
  content <- xml_text(rev_nodes)
  
  # Return data frame 
  data.frame(user = user,
    timestamp = timestamp,
    content = substr(content, 1, 40))
}

# Call function for "Hadley Wickham"
get_revision_history("Hadley Wickham")

```

Nice job! Your function parsed the XML data, but you could have just as easily parsed the JSON data.

