---
title: "Downloading Static Formatted Files"
author: "Jane Kathambi"
date: "11 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---
# Introduction

We are goig to explore how to download and read in static files.

# Load the required libraries

```{r}
library(dplyr)

```

# Downloading files from the internet and reading them into R

We're going to look at reading already-formatted datasets - CSV or TSV files - into R from the internet. 

This is a lot easier than it might sound because R's file-reading functions accept not just file paths, but also URLs.



# Without saving raw files to a disk

We simply assign the read data to a variable

```{r}
# Here are the URLs! As you can see they're just normal strings
csv_url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1561/datasets/chickwts.csv"
tsv_url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_3026/datasets/tsv_data.tsv"

# Read a file in from the CSV URL and assign it to csv_data
csv_data <- read.csv(csv_url)

# Read a file in from the TSV URL and assign it to tsv_data
tsv_data <- read.delim(tsv_url)
```

# Saving raw files to a disk

Sometimes just reading the file in from the web is enough, but often you'll want to store it locally so that you can refer back to it. 

This also lets you avoid having to spend the start of every analysis session twiddling your thumbs while particularly large files download.

Helpfully, R has _**download.file()**_, a function that lets you do just that: download a file to a location of your choice on your computer. 

It takes two arguments; _**url**_, indicating the URL to read from, and _**destfile**_, the destination to write the downloaded file to. 

After downolading we then read the data into R. In this case we  will use read.csv() function.

In this case, we'll use the pre-defined URL - csv_url

```{r}
# Download the file with download.file()
download.file(url = csv_url, destfile ="feed_data.csv" )

# Read it in with read.csv()
csv_data <- read.csv("feed_data.csv")
```

# Saving formatted files to disk

Whether we're downloading the raw files with download.file() or using read.csv() and its sibling functions, at some point we're probably going to find the need to modify our input data, and then save the modified data to disk so that we don't lose the changes.

We could use write.table(), but then we have to worry about accidentally writing out data in a format R can't read back in. 

An easy way to avoid this risk is to use saveRDS() and readRDS(), which save R objects in an R-specific file format, with the data structure intact. 

That means we can use it for any type of R object (even ones that don't turn into tables easily), and not worry we'll lose data reading it back in.

saveRDS() takes two arguments, **object**, pointing to the R object to save and **file** pointing to where to save it to. NB: Unlike save() saveRDS() does not save the name of the object to the disk but it aonly saves the representation of the object. This means that when we read the data back into r we can assign it a name of our choice.

readRDS() expects **file**, referring to the *path* to the RDS file to read in.

In this example we're going to modify the data we already read in, which is predefined as csv_data, and write the modified version out to a file before reading it in again.

```{r}
# Add a new column: square_weight
csv_data$square_weight <- csv_data$weight^2

# Save it to disk with saveRDS()
saveRDS(object = csv_data, file ="modified_feed_data.RDS")

# Read it back in with readRDS()
modified_feed_data <- readRDS(file ="modified_feed_data.RDS")

# Examine modified_feed_data
glimpse(modified_feed_data)
```
