---
title: "Respectful API Usage"
author: "Jane Kathambi"
date: "11 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Introduction

There are two aspects of respectful API usage:
1. User Agents
2. Rate Limiting

# Load the required libraries

```{r}
library(httr)
```

# User agents

* Bits of text that ID your browser (or software)
* Gives the server some idea of what you're trying to do
* You can set one with your requests with user_agent()
* Add an email address so they can contact you.

Informative user-agents are a good way of being respectful of the developers running the API you're interacting with. They make it easy for them to contact you in the event something goes wrong. I always try to include:

1. My email address;
2. A URL for the project the code is a part of, if it's got a URL.

Building user agents is done by passing a call to user_agent() into the GET() or POST() request; something like:

GET("http://url.goes.here/", user_agent("somefakeemail@domain.com http://project.website"))

In the event you don't have a website, a short one-sentence description of what the project is about serves pretty well.

```{r}
# Do not change the url
url <- "https://wikimedia.org/api/rest_v1/metrics/pageviews/per-article/en.wikipedia/all-access/all-agents/Aaron_Halfaker/daily/2015100100/2015103100"

# Add the email address and the test sentence inside user_agent()
server_response <- GET(url, user_agent("my@email.address this is a test"))
```

 From your end, the request looks exactly the same with or without a user agent, but for the server the difference can be vital.

# Rate limiting

* Too many requests makes for a sad server
* Deliberately slows down your code to keep under a desired 'rate'
* Slows you, but avoids getting you banned from the server

Rate-limiting: making sure you only make a certain number of requests to the server in a given time period. 

Your limit will vary from server to server, but the implementation is always pretty much the same and involves a call to Sys.sleep(). This function takes one argument, a number, which represents the number of seconds to "sleep" (pause) the R session for. So if you call Sys.sleep(15), it'll pause for 15 seconds before allowing further code to run.

As you can imagine, this is really useful for rate-limiting. If you are only allowed 4 requests a minute? No problem! Just pause for 15 seconds between each request and you're guaranteed to never exceed it. Let's demonstrate how by putting together a little loop that sends multiple requests, on a 5-second time delay.

```{r}
# Construct a vector of 2 URLs
urls <- c("http://fakeurl.com/api/1.0/", "http://fakeurl.com/api/2.0/")

for(url in urls){
    # Send a GET request to url
    result <- GET(url)
    # Delay for 5 seconds between requests
    Sys.sleep(5)
}
```

With this and the user agents we covered in the last exercise, you know enough to avoid annoying servers - and thus snarling up your code.