---
title: "Tying it all together"
author: "Jane Kathambi"
date: "11 July 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Tying it all together
Using everything that you learned in the chapter, let's make a simple replica of one of the 'pageviews' functions - building queries, sending GET requests (with an appropriate user agent) and handling the output in a fault-tolerant way.

You'll build this function up piece by piece in this exercise.

To output an error, you will use the function stop(), which takes a string as an argument, stops the execution of the program, and outputs the string as an error.

You can try it right now by running stop("This is an error").

# Load the required libraries

```{r}
library(httr)
```

# First, get the function to construct the url.

We will Construct a query: Directory based API urls.

In the call to paste(), add article_title as the second argument to construct url

```{r}
get_pageviews <- function(article_title){
  
  # construct the url
  
  url <- paste(
    "https://wikimedia.org/api/rest_v1/metrics/pageviews/per-article/en.wikipedia/all-access/all-agents", 
    
    # Include article_title
    article_title, 
    
    "daily/2015100100/2015103100", 
    
    sep = "/"
  ) 
  
  # Return the url
  url
}
```

# Now, make the request.

Use GET() to request url with a user agent "my@email.com this is a test"

```{r}
get_pageviews <- function(article_title){
  
  # construct the url
  
  url <- paste(
    "https://wikimedia.org/api/rest_v1/metrics/pageviews/per-article/en.wikipedia/all-access/all-agents", 
    
    article_title, 
    
    "daily/2015100100/2015103100", 
    
    sep = "/"
  ) 
  
  # Get the webpage  
  response <- GET(url, config = user_agent("my@email.com this is a test")) 
  
  # Return the response
  response
}
```

# Now, add an error check.

Check the response for errors with http_error(), throwing an error of "the request failed" with stop() if there was one.

```{r}
get_pageviews <- function(article_title){
  
  # construct the url
  
  url <- paste(
    "https://wikimedia.org/api/rest_v1/metrics/pageviews/per-article/en.wikipedia/all-access/all-agents", 
    
    article_title, 
    
    "daily/2015100100/2015103100", 
    
    sep = "/"
  ) 
  
  
  # Get the webpage  
  
  response <- GET(url, config = user_agent("my@email.com this is a test")) 
  
  # Is there an HTTP error?
  if(http_error(response)){ 
    # Throw an R error
     stop("the request failed") 
  }
  
  # Return the response
  response
}
```

# Return content of response instead of response

Finally, instead of returning response, return the content() of the response.

```{r}
get_pageviews <- function(article_title){
  
  # construct the url
  
  url <- paste(
    "https://wikimedia.org/api/rest_v1/metrics/pageviews/per-article/en.wikipedia/all-access/all-agents", 
    
    article_title, 
    
    "daily/2015100100/2015103100", 
    
    sep = "/"
  ) 
  
  
  # Get the webpage  
  
  response <- GET(url, config = user_agent("my@email.com this is a test")) 
  
  # Is there an HTTP error?
  if(http_error(response)){ 
    # Throw an R error
     stop("the request failed") 
  }
  
  # Return the response's content
  content(response)
}
```
