---
title: "Classification: Ridge Penalised Logistic Regression Diabetes Data"
author: "Jane Kathambi"
date: "8 June 2018"
output: 
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Introduction
A classification problem involves predicting a non-numerical value—that is, a categorical variable, also known as discrete variable.

Most of the classification algorithms computes the probability of belonging to a given class. Observations are then assigned to the class that have the highest probability score.

Generally, you need to decide a probability cutoff above which you consider the an observation as belonging to a given class.

## Recap:
Use Logistic Regression with ridge penalty when the classes are only two and are inseparable.
Use radial kernel svms for inseparable data. Can be used for two or more than two classes. FDA (FDA is a flexible extension of LDA that uses non-linear combinations of predictors such as splines. FDA is useful to model multivariate non-normality or non-linear relationships among variables within each group, allowing for a more accurate classification.) works well too with any number of classes. 

So for inseparabe data use Radial kernel svm, FDA or Logistic Regression with ridge penalty.

Generally, the purpose of regularization is to balance accuracy and simplicity.
This means a model with the smallest number of predictors that also gives a good accuracy. 

# The PimaIndiansDiabetes2 Data
PimaIndiansDiabetes2 is contained in the mlbench package. We will use it for predicting the probability of being diabetes positive or negative based on multiple clinical variables. This is a binary classification problem.

This data set has two inseparable classes as we will see and so logistic regression will model this data well.

The data contains 768 individuals (female) and 9 clinical variables for predicting the probability of individuals in being diabete-positive or negative:

* pregnant: number of times pregnant
* glucose: plasma glucose concentration
* pressure: diastolic blood pressure (mm Hg)
* triceps: triceps skin fold thickness (mm)
* insulin: 2-Hour serum insulin (mu U/ml)
* mass: body mass index (weight in kg/(height in m)^2)
* pedigree: diabetes pedigree function
* age: age (years)
* diabetes: class variable

#Load the required libraries.
* glmnet package. Carries the ridge, lasso, and elsatic net regression function.
* tidyverse. For data mangling.
* caret. createDataPartition for partitioning the data into test set and train set.

```{r}
library(glmnet)
library(tidyverse)
library(caret)

```

# Exploring the PimaIndiansDiabetes2 Data

```{r}
# Load the data
data("PimaIndiansDiabetes2", package = "mlbench")

# remove Nas
PimaIndiansDiabetes2 <- na.omit(PimaIndiansDiabetes2)

# Inspect the data
sample_n(PimaIndiansDiabetes2, 3)

# remove the first column since it is not a feature
PimaIndiansDiabetes2<-PimaIndiansDiabetes2[, -1]

# Covariance scatterplot Matrix. View how each variable varies with the rest as well as how the classes are distributed.
pairs(PimaIndiansDiabetes2, col=PimaIndiansDiabetes2$diabetes)

# plot two variables to see how many classes there are and if they are separable.
PimaIndiansDiabetes3=PimaIndiansDiabetes2[,c('glucose','pressure')]
pairs(PimaIndiansDiabetes3, col=PimaIndiansDiabetes2$diabetes)

# dim
dim(PimaIndiansDiabetes2)
```

From the pairs plot above we see that there are two classes which are inseparable. So logistic regression will work well in this case.Radial kernel svm would work well too.

## Normalize the data. Categorical variables are automatically ignored.
```{r}
# Estimate preprocessing parameters
preproc.param <- PimaIndiansDiabetes2 %>% 
  preProcess(method = c("center", "scale"))

# Transform the data using the estimated parameters
PimaIndiansDiabetes2_normalized <- preproc.param %>% predict(PimaIndiansDiabetes2)

```


# Split the data into training and test set
We'll randomly split the data into training set (80% for building a predictive model) and test set (20% for evaluating the model). Make sure to set seed for reproductibility.
```{r}
set.seed(123)
training.samples <- PimaIndiansDiabetes2_normalized$diabetes %>% 
  createDataPartition(p = 0.8, list = FALSE)

train.data  <- PimaIndiansDiabetes2_normalized[training.samples, ]
test.data <- PimaIndiansDiabetes2_normalized[-training.samples, ]

```

# Fitting ridge penalized logistic regression model
## Additionnal data preparation
The R function model.matrix() helps to create the matrix of predictors and also automatically converts categorical predictors to appropriate dummy variables, which is required for the glmnet() function.
```{r}
# Dumy code categorical predictor variables
x.train <- model.matrix(diabetes~., train.data)

# Convert the outcome (class) to a numerical variable
y.train <- ifelse(train.data$diabetes == "pos", 1, 0)

# test data
x.test <- model.matrix(diabetes ~., test.data)
```

To fit a ridge penalised logistic regression we use the glmnet package and set alpha=0.

glmnet(x,y, family="binomial", alpha=0, lamda=NULL)

* x: matrix of predictor variables.
* y: the repsonse variable which is a binary variable.
* family: response type, use "binomial" for a binary outcome.
* alpha: the elasticnet mixing parameter. Allowed values include:
    + 1: for lasso regression.
    + 0: for ridge regression.
    + A value between 0 and 1 (say 0.3) for elastic net regression.
* lambda: a numeric value defining the amount of shrinkage. The analyst should specify. The best lambda is the one which minimizes the cross-validation prediction error rate. This can be determined automatically using cv.glmnet(). than the one obtained with lambda.min.
Setting lamda=lambda.1se produces a simpler model(less variables) as compared to lambda.min, but the model might be a little bit less accurate.

so:
1. Find the best lambda using cross-validation. 
    + cv.ridge<-cv.glmnet(x,y,alpha=0, family="binomial")
    + best_lambda<-cv.ridge$lambda.min
    
2.  Use this lambda to fit the model on the training data.
    + fit.ridge<-glmnet(x,y,alpha=0, family="binomial", lambda=best_lambda)


```{r}
# Find the best lambda using cross-validation
set.seed(123) 
cv.lasso <- cv.glmnet(x.train, y.train, alpha = 0, family = "binomial")# binomial means logistic regression and alpha=0 means ridge regression, so ridge penalised logistic regression.

# Fit the final model on the training data using best lambda
model <- glmnet(x.train, y.train, alpha = 0, family = "binomial",
                lambda = cv.lasso$lambda.min)

# Display regression coefficients
coef(model)

# Make predictions on the test data
probabilities <- model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "pos", "neg")

# Model accuracy
observed.classes <- test.data$diabetes
mean(predicted.classes == observed.classes)
```

Our model has an accuracy of 0.7820 which implies that the model is doing fairly well.
The best model for this data is fda.

# Model Accuracies Ranking
The following is a ranking of how the models perfomed on classifying this diabetes data strating from the best model. The best model for this data is FDA.

1. Flexible Discriminant Analysis: 0.807 accuracy
2. Radial Kernel SVM classifier: 0.794 accuracy
3. Ridge Penalised Logistic Regression: 0.782 accuracy
4. Mixture Discriminant Analysis: 0.769 accuracy


