---
title: "Classification Ridge Penalised Logistic Regression. Smarket Data set"
author: "Jane Kathambi"
date: "8 June 2018"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---
# Introduction
A classification problem involves predicting a non-numerical value—that is, a categorical variable, also known as discrete variable.

Most of the classification algorithms computes the probability of belonging to a given class. Observations are then assigned to the class that have the highest probability score.

Generally, you need to decide a probability cutoff above which you consider the an observation as belonging to a given class.

## Recap:
Use Logistic Regression with ridge penalty when the classes are only two and are inseparable.
Use radial kernel svms for inseparable data. Can be used for two or more than two classes. FDA (FDA is a flexible extension of LDA that uses non-linear combinations of predictors such as splines. FDA is useful to model multivariate non-normality or non-linear relationships among variables within each group, allowing for a more accurate classification.) works well too with any number of classes. 

So for inseparabe data use Radial kernel svm, FDA or Logistic Regression with ridge penalty.

Generally, the purpose of regularization is to balance accuracy and simplicity.
This means a model with the smallest number of predictors that also gives a good accuracy. 

# The Stock Market Data
The Smarket data, is part of the ISLR library. 

This data set consists of percentage returns for the S&P 500 stock index over 1250 days, from the beginning of 2001 until the end of 2005. 

Each row is a date representing today, and for each date(today), the percentage returns for each of the five previous trading days, Lag1 through Lag5, have been recorded. Year, volume , and direction have been recorded too for each date(today. 

The various variables are listed below:

* Year
    + The year that the observation was recorded
* Lag1
    + Percentage return for previous day
* Lag2
    + Percentage return for 2 days previous
* Lag3
    + Percentage return for 3 days previous
* Lag4
    + Percentage return for 4 days previous
* Lag5
    + Percentage return for 5 days previous
* Volume
    + Volume of shares traded (number of daily shares traded in billions)
* Today
    + Percentage return for today (date in question)
* Direction
    + A factor with levels Down and Up indicating whether the market had a positive or negative return on a given day
    
The Smarket data has two inseparable classes as we will see shortly. So logistic regression will model this data well.

#Load the required libraries.
* ISLR package. Carries the Smarket dataset.
* * glmnet package. Carries the ridge, lasso, and elsatic net regression functions.
* tidyverse. For data mangling.
* caret. createDataPartition for partitioning the data into test set and train set.Preprocess for data preprocessing.

```{r}
library(caret)
library(ISLR)
library(tidyverse)
library(glmnet)
```

# Exploring the Stock Market Data
We will begin by examining some numerical and graphical summaries of the Smarket data. 

```{r}
dim(Smarket)

names(Smarket)

summary(Smarket)

glimpse(Smarket)

head(Smarket)

tail(Smarket)

# Covariance scatterplot Matrix. View how each variable varies with the rest as well as how the classes are distributed.
pairs(Smarket, col=Smarket$Direction)

# plot two variables to see how many classes there are and if they are separable.
Smarket2=Smarket[,c('Lag1','Lag2')]
pairs(Smarket2, col=Smarket$Direction)
```

It is evident from the covariance matix that this data set has two classes that are inseparable. So logistic regression will model this data well.

# Preparing the data
It is always good to standardize the data before using it.

## Normalize the data. Categorical variables are automatically ignored.

```{r}
# Estimate preprocessing parameters
preproc.param <- Smarket %>% 
  preProcess(method = c("center", "scale"))

# Transform the data using the estimated parameters
standardized_smarket <- preproc.param %>% predict(Smarket)

```

## Split the data into training and test set
```{r}
# Split the data into training (80%) and test set (20%)
set.seed(123)
training.samples <- standardized_smarket$Direction %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data <- standardized_smarket[training.samples, ]
test.data <- standardized_smarket[-training.samples, ]
```

## Additionnal data preparation
The R function model.matrix() helps to create the matrix of predictors and also automatically converts categorical predictors to appropriate dummy variables, which is required for the glmnet() function.

```{r}
# Matrix of predictor variables
x.train <- model.matrix(Direction~., train.data)

# Convert the outcome (Direction) to a numerical variable
y.train <- ifelse(train.data$Direction == "Up", 1, 0)

# test data
x.test <- model.matrix(Direction ~., test.data)
```

# Fitting ridge penalized logistic regression model
To fit a ridge penalised logistic regression we use the glmnet package and set alpha=0.

glmnet(x,y, family="binomial", alpha=0, lamda=NULL)

* x: matrix of predictor variables.
* y: the repsonse variable which is a binary variable.
* family: response type, use "binomial" for a binary outcome.
* alpha: the elasticnet mixing parameter. Allowed values include:
    + 1: for lasso regression.
    + 0: for ridge regression.
    + A value between 0 and 1 (say 0.3) for elastic net regression.
* lambda: a numeric value defining the amount of shrinkage. The analyst should specify. The best lambda is the one which minimizes the cross-validation prediction error rate. This can be determined automatically using cv.glmnet(). than the one obtained with lambda.min.
Setting lamda=lambda.1se produces a simpler model(less variables) as compared to lambda.min, but the model might be a little bit less accurate.

so:
1. Find the best lambda using cross-validation. 
    + cv.ridge<-cv.glmnet(x,y,alpha=0, family="binomial")
    + best_lambda<-cv.ridge$lambda.min
    
2.  Use this lambda to fit the model on the training data.
    + fit.ridge<-glmnet(x,y,alpha=0, family="binomial", lambda=best_lambda)

## Find the best lambda

```{r}
# Find the best lambda using cross-validation
set.seed(123) 
cv.lasso <- cv.glmnet(x.train, y.train, alpha = 0, family = "binomial")# binomial means logistic regression and alpha=0 means ridge regression, so ridge penalised logistic regression.
```

## Fit the model using the best lambda

```{r}
# Fit the final model on the training data using best lambda
model <- glmnet(x.train, y.train, alpha = 0, family = "binomial",
                lambda = cv.lasso$lambda.min)

# Display regression coefficients
coef(model)
```

## Accessing model accuracy

```{r}

# Make predictions on the test data
probabilities <- model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Up", "Down")

# Model accuracy
observed.classes <- test.data$Direction
mean(predicted.classes == observed.classes)
```

Our model has an accuracy of 0.8835 which implies that the model is doing fairly well.

The best model for this data is radial kernel svm classifier.

# Model Accuracies Ranking
The following is a ranking of how the models perfomed on classifying this smarket data starting from the best model. 
1. Radial Kernel SVM classifier:  0.9518 accuracy
2. Flexible Discriminant Analysis: 0.9438 accuracy
3. Mixture Discriminant Analysis: 0.9157 accuracy
4. Ridge Penalised Logistic Regression: 0.8835 accuracy

