---
title: "Classification Flexible Discriminant Analysis. Diabetes Data"
author: "Jane Kathambi"
date: "8 June 2018"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---
# Introduction
A classification problem involves predicting a non-numerical value—that is, a categorical variable, also known as discrete variable.

Most of the classification algorithms computes the probability of belonging to a given class. Observations are then assigned to the class that have the highest probability score.

Generally, you need to decide a probability cutoff above which you consider the an observation as belonging to a given class.

## Recap
** separable data:** 
* LDA: Small training set. Also Linear kernel svms.
* QDA: Large training set. Also Linear kernel svms.
* RDA: Large training set and too many features. Also Linear kernel svms.

**inseparable data:**
* FDA: Multivariate inseparable data sets. Also radial kernel svms.
* MDA: Classes have sub-classes which are distributed. Also radial kernel svms.

# The PimaIndiansDiabetes2 Data
PimaIndiansDiabetes2 is contained in the mlbench package. We will use it for predicting the probability of being diabetes positive or negative based on multiple clinical variables. This is a binary classification problem.

This data set has two inseparable classes, as we will see shortly, and so an fda will model this data well.

The data contains 768 individuals (female) and 9 clinical variables for predicting the probability of individuals in being diabete-positive or negative:

* pregnant: number of times pregnant
* glucose: plasma glucose concentration
* pressure: diastolic blood pressure (mm Hg)
* triceps: triceps skin fold thickness (mm)
* insulin: 2-Hour serum insulin (mu U/ml)
* mass: body mass index (weight in kg/(height in m)^2)
* pedigree: diabetes pedigree function
* age: age (years)
* diabetes: class variable

# Classification:  FDA (Flexible discriminant analysis)
FDA is a flexible extension of LDA that uses non-linear combinations of predictors such as splines. FDA is useful to model multivariate non-normality or non-linear relationships among variables within each group, allowing for a more accurate classification.


# Load the required libraries.
* tidyverse. for easy data manipulation and visualization
* caret. for easy machine learning workflow. createDataPartition for partitioning the data into test set and train set.
* fda library: has the mda function.


```{r}
library(tidyverse)
library(caret)
library(fda)
```

# Exploring the PimaIndiansDiabetes2 Data

```{r}
# Load the data
data("PimaIndiansDiabetes2", package = "mlbench")

# remove Nas
PimaIndiansDiabetes2 <- na.omit(PimaIndiansDiabetes2)

# Inspect the data
sample_n(PimaIndiansDiabetes2, 3)

# Covariance scatterplot Matrix. View how each variable varies with the rest as well as how the classes are distributed.
pairs(PimaIndiansDiabetes2, col=PimaIndiansDiabetes2$diabetes)

# plot two variables to see how many classes there are and if they are separable.
PimaIndiansDiabetes3=PimaIndiansDiabetes2[,c('glucose','pressure')]
pairs(PimaIndiansDiabetes3, col=PimaIndiansDiabetes2$diabetes)

# dim
dim(PimaIndiansDiabetes2)
```

From the pairs plot above we see that there are two classes which are inseparable. So fda will model this data well. Radial kernel svm and ridge penalised logistic regression would too work well.

## Normalize the data. Categorical variables are automatically ignored.
```{r}
# Estimate preprocessing parameters
preproc.param <- PimaIndiansDiabetes2 %>% 
  preProcess(method = c("center", "scale"))

# Transform the data using the estimated parameters
PimaIndiansDiabetes2_normalized <- preproc.param %>% predict(PimaIndiansDiabetes2)

```


## Split the data into training and test set
We'll randomly split the data into training set (80% for building a predictive model) and test set (20% for evaluating the model). Make sure to set seed for reproductibility.
```{r}
set.seed(123)
training.samples <- PimaIndiansDiabetes2_normalized$diabetes %>% 
  createDataPartition(p = 0.8, list = FALSE)

train.data  <- PimaIndiansDiabetes2_normalized[training.samples, ]
test.data <- PimaIndiansDiabetes2_normalized[-training.samples, ]

```


# Fit the FDA model
```{r}
# Fit the FDA model
fda_model <- fda(diabetes~., data = train.data)
fda_model
```

# FDA Model accuracy
For fda model the output of predictions is predicted classes. So we access model accuracy as follows:
mean(preds == test.data$diabetes)
```{r}
# Make predictions
preds <- fda_model %>% predict(test.data)
# Model accuracy
mean(preds == test.data$diabetes)


```

Our model has an accuracy of 0.8076. This implies that the model is doing fairly well. FDA is the best model for this data.

# Model Accuracies Ranking
The following is a ranking of how the models perfomed on classifying this diabetes data strating from the best model. The best model for this data is FDA.

1. Flexible Discriminant Analysis: 0.807 accuracy
2. Radial Kernel SVM classifier: 0.794 accuracy
3. Ridge Penalised Logistic Regression: 0.782 accuracy
4. Mixture Discriminant Analysis: 0.769 accuracy
